---
title: "EDA to Guide Feature Engineering"
output: html_notebook
---

## Ingests Data
This first chunk just pulls the data we ingested into the local environment.

```{r Pulls Data In}
library(magrittr)
data <- readr::read_rds(here::here("./data/tidy/tidy-data.rds"))
```

## Summarizes Dataset
```{r Summary}
psych::describe(data)
```

## Potential Outcome Variables
Our outcome variables (as indicated in the [original study](https://repositorio.inesctec.pt/bitstream/123456789/7179/1/P-00N-9XR.pdf) and a [subsequent study](https://ieeexplore.ieee.org/document/8070120) are the four types of tests that one uses to diagnose HPV:

* Hinselmann
* Schiller
* Cytology
* Biospy

```{r Cancer Diagnoses}
data %>% 
  dplyr::summarise(
    hinselmann_pos = sum(hinselmann),
    schiller_pos = sum(schiller),
    cytology_pos = sum(citology),
    biopsy_pos = sum(biopsy)
    )
```

```{r all tests pos}
data %>%
  dplyr::filter(
    hinselmann == 1 &
      schiller == 1 &
      citology == 1 &
      biopsy == 1
    ) %>%
  dplyr::tally()
```

Only 6 records have all four show positive.

```{r any tests pos}
data %>%
  dplyr::filter(
    hinselmann == 1 |
      schiller == 1 |
      citology == 1 |
      biopsy == 1
    ) %>%
  dplyr::tally()
```

However, 102 show at least one of these tests being positive.

## Potential Risk Factors
I looked at the following sources to consider potential risk factors:

 * The [CDC](https://www.cdc.gov/cancer/cervical/basic_info/risk_factors.htm)
 * The [American Cancer Society](https://www.cancer.org/cancer/cervical-cancer/causes-risks-prevention/risk-factors.html)
 * [Cancer Research UK](https://www.cancerresearchuk.org/about-cancer/cervical-cancer/risks-causes)
 
All of them seem to suggest the following as risk factors: 

* Having HIV
* Smoking
* Using birth control pills for a long time (five or more years)
* Having given birth to three or more children
* Having several sexual partners
* Sexual history at a young age (before 18)
* Having chlamydia or other STIs
* Previous cancer diagnoses

They also recognize the following as a potentially protective factor:

* Using an IUD

### Creating Boolean Indicators to Examine Those Factors

A few of our risk factors need to be turned into a boolean factor, particularly age of first sexual activity (under 18), number of pregancies (3+), number of sexual partners (3+), and birth control pills (5+) before we can use them later. 

```{r Risk Factors}
analytical_set <- data %>%
  dplyr::mutate(
    underage_sex = ifelse(first_sexual_intercourse < 18, 1,
                          ifelse(is.na(first_sexual_intercourse), NA, 0)),
    many_preg = ifelse (num_of_pregnancies >= 3, 1,
                        ifelse(is.na(num_of_pregnancies), NA, 0)),
    many_sp = ifelse (number_of_sexual_partners >= 3, 1,
                        ifelse(is.na(number_of_sexual_partners), NA, 0)),
    long_bc = ifelse (hormonal_contraceptives_years >= 5, 1,
                        ifelse(is.na(hormonal_contraceptives_years), NA, 0))) %>%
  dplyr::select(-c(
    age, number_of_sexual_partners, first_sexual_intercourse, 
    num_of_pregnancies, smokes_years, smokes_packs_year, 
    hormonal_contraceptives_years, iud_years, stds_number,
    stds_number_of_diagnosis, stds_time_since_first_diagnosis,
    stds_time_since_first_diagnosis)) %>%
  dplyr::select(hinselmann, schiller, citology, biopsy, dplyr::everything())
```

```{r Risk Factors Present}
analytical_set %>%
    dplyr::summarize_if(is.numeric, sum, na.rm=TRUE)
```

Several of the stds may need to be excluded in the analytical file because they'll mathematically drop out anyway with so few cases, but let's examine the counts by each outcome variable.

```{r Risk Factors Hinselmann}
analytical_set %>%
  dplyr::filter(hinselmann == 1) %>%
  dplyr::select(-c(hinselmann, schiller, citology, biopsy)) %>%
    dplyr::summarize_if(is.numeric, sum, na.rm=TRUE)
```

```{r Risk Factors Schiller}
analytical_set %>%
  dplyr::filter(schiller == 1) %>%
  dplyr::select(-c(hinselmann, schiller, citology, biopsy)) %>%
    dplyr::summarize_if(is.numeric, sum, na.rm=TRUE)
```

```{r Risk Factors Cytology}
analytical_set %>%
  dplyr::filter(citology == 1) %>%
  dplyr::select(-c(hinselmann, schiller, citology, biopsy)) %>%
    dplyr::summarize_if(is.numeric, sum, na.rm=TRUE)
```

```{r Risk Factors Biopsy}
analytical_set %>%
  dplyr::filter(biopsy == 1) %>%
  dplyr::select(-c(hinselmann, schiller, citology, biopsy)) %>%
    dplyr::summarize_if(is.numeric, sum, na.rm=TRUE)
```

```{r Risk Factors ANY}
analytical_set %>%
  dplyr::filter(hinselmann == 1 |
                  schiller == 1 |
                  citology == 1 |
                  biopsy == 1) %>%
  dplyr::select(-c(hinselmann, schiller, citology, biopsy)) %>%
    dplyr::summarize_if(is.numeric, sum, na.rm=TRUE)
```

This may help significantly in being able to calculate a "risk score" for testing positive for cervical cancer using any of these risk / protective factors.

```{r clean up}
rm(list = ls())
```