---
title: "Ingest & Tidy Data"
output: html_notebook
---

## Sourced Functions
This sources all the functions we will need to ingest and tidy our data!
```{r Sourced Functions}
source(here::here("/src/ingest.R"))
```

## Ingest
This block ingests the Cervical cancer (Risk Factors) data from [UCI's Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Cervical+cancer+%28Risk+Factors%29). The data are not likely to change, so only run this once!

```{r Ingest Data, eval=FALSE}
getdata("https://archive.ics.uci.edu/ml/machine-learning-databases/00383/risk_factors_cervical_cancer.csv")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

