---
title: "Ingest & Tidy Data"
output: html_notebook
---

## Sourced Functions
This sources all the functions we will need to ingest and tidy our data!
```{r Sourced Functions}
library(magrittr)
source(here::here("./r/ingest.R"))
source(here::here("./r/coltype.R"))
source(here::here("./r/removechar.R"))
source(here::here("./r/colclean.R"))
```

## Ingest
This block ingests the Cervical Cancer (Risk Factors) Data from [UCI's Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Cervical+cancer+%28Risk+Factors%29). The data are not likely to change, so only run this once!

```{r Ingest Data}
ingest("https://archive.ics.uci.edu/ml/machine-learning-databases/00383/risk_factors_cervical_cancer.csv")
```

##Tidy
This first chunk just pulls the data we ingested into the local environment.

```{r Pulls Data In}
data <- readr::read_rds(here::here("./data/raw/raw-data.rds"))
```

This next chunk fixes the instances of "?" in the data. It first changes all columns to characters, replaces the "?" with NA, and then converts the columns to numeric. The function "coltype" takes a dataframe (x) and converts all column types to the type specified (y). The function "removechar" it uses takes any dataset (x), value that a cell might have (y), and replaces it with any other value (z). 

```{r Cleans ? Marks}
data <- coltype(data, as.character)
data <- removechar(data, "?", NA)
data <- coltype(data, as.numeric)
```

This next chunk fixes the column names to be all lower case and with underscores between words. The function "colclean" takes a dataset (x) and replaces all instances of a character (y) with a new character (z).

```{r Cleans Column Names}
data <- colclean(data, "\\.$", "")
data <- colclean(data, "\\.\\.", "_")
data <- colclean(data, "\\.", "_")
colnames(data) <- tolower(colnames(data))
```

This next section saves the data to a tidy data frame & cleans the environment.
```{r Saves Data}
data %>% readr::write_rds(here::here("./data/tidy/tidy-data.rds"))
rm(list = ls())
```